---
title: The categorize function
author: "Umut Caglar"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
  
# The categorize function
This is a document provide examples related with categorize function. It will investigate 2 examples one is related with sigmoidal and other is related with double sigmoidal
  
  
```{r install packages, echo=FALSE, warning=FALSE, results='hide',message=FALSE}

###*****************************
# INITIAL COMMANDS TO RESET THE SYSTEM
rm(list = ls())
if (is.integer(dev.list())){dev.off()}
cat("\014")
seedNo=14159
set.seed(seedNo)
###*****************************

###*****************************
require("sicegar")
require("dplyr")
require("ggplot2")
###*****************************
```

## Data generation
To generate the figures, we will go backwards and firstly generate some data to analize. To add some randomness to the input data I will use some noise. The input of all package must be in the form of a data frame with at least 2 columns time and intensity. 

We will generate data for both sigmoidal and double sigmoidal 

### Generate data for sigmoidal
```{r generate data for sigmoidal}
time=seq(3,24,0.5)

#simulate intensity data and add noise
noise_parameter=0.1
intensity_noise=stats::runif(n = length(time),min = 0,max = 1)*noise_parameter
intensity=sigmoidalFitFormula(time, maximum=4, slopeParam=1, midPoint=8)
intensity=intensity+intensity_noise

dataInputSigmoidal=data.frame(intensity=intensity,time=time)
```


### Generate data for double - sigmoidal
```{r generate data for double - sigmoidal}
noise_parameter=0.2
intensity_noise=runif(n = length(time),min = 0,max = 1)*noise_parameter
intensity=doublesigmoidalFitFormula(time,
                                    finalAsymptoteIntensityRatio=.3,
                                    maximum=4,
                                    slope1Param=1,
                                    midPoint1Param=7,
                                    slope2Param=1,
                                    midPointDistanceParam=8)
intensity=intensity+intensity_noise

dataInputDoubleSigmoidal=data.frame(intensity=intensity,time=time)
```

## Data normalization 

This is the first step. Data should be normalized before any fit. I.e time and intensity should be in between 0-1 interval. 

The normalization code is 

```{r normalize_data}
normalizedSigmoidalInput = sicegar::normalizeData(dataInput = dataInputSigmoidal, 
                                         dataInputName = "sigmoidalSample")

normalizedDoubleSigmoidalInput = sicegar::normalizeData(dataInput = dataInputDoubleSigmoidal, 
                                         dataInputName = "doubleSigmoidalSample")
```


## Fits for sigmoidal and double sigmoidal

Now we need to fit all three models to the data. Linear, sigmoidal and double sigmoidal on both sigmoidal and double-sigmoidal data.


### Fit Linear, Sigmoidal and Double-Sigmoidal models to Sigmoidal data
```{r linear sigmoidal and double-sigmoidal fits to sigmoidal data}
# Do the sigmoidal fit
# Fit linear model
linearModel_sd=fitFunction(dataInput=normalizedSigmoidalInput,
                           model="linear",
                           n_runs_min=20,
                           n_runs_max=500,
                           showDetails=FALSE)

# Fit sigmoidal model
sigmoidalModel_sd=fitFunction(dataInput=normalizedSigmoidalInput,
                              model="sigmoidal",
                              n_runs_min=20,
                              n_runs_max=500,
                              showDetails=FALSE)

# Fit double sigmoidal model
doubleSigmoidalModel_sd=fitFunction(dataInput=normalizedSigmoidalInput,
                                    model="doublesigmoidal",
                                    n_runs_min=20,
                                    n_runs_max=500,
                                    showDetails=FALSE)
```


### Fit Linear, Sigmoidal and Double-Sigmoidal models to Double-Sigmoidal data
```{r linear sigmoidal amd double-sigmoidal fits to double-sigmoidal data}
# Do the sigmoidal fit
# Fit linear model
linearModel_dsd=fitFunction(dataInput=normalizedDoubleSigmoidalInput,
                            model="linear",
                            n_runs_min=20,
                            n_runs_max=500,
                            showDetails=FALSE)

# Fit sigmoidal model
sigmoidalModel_dsd=fitFunction(dataInput=normalizedDoubleSigmoidalInput,
                               model="sigmoidal",
                               n_runs_min=20,
                               n_runs_max=500,
                               showDetails=FALSE)

# Fit double sigmoidal model
doubleSigmoidalModel_dsd=fitFunction(dataInput=normalizedDoubleSigmoidalInput,
                                     model="doublesigmoidal",
                                     n_runs_min=20,
                                     n_runs_max=500,
                                     showDetails=FALSE)
```

## Model decision

Now we have attempts of fits for linear, sigmoidal and double sigmoidal models on our datasets. By using the results of those atempts one can make decisions about the category of the input data; wheather it is sigmoidal or double sigmoidal.

```{r decide wheather the data is sigmoidal or double sigmoidal}
outputCluster_sd=categorize(parameterVectorLinear=linearModel_sd,
                            parameterVectorSigmoidal=sigmoidalModel_sd,
                            parameterVectorDoubleSigmoidal=doubleSigmoidalModel_sd)

print(outputCluster_sd) # This should give sigmoidal 

outputCluster_dsd=categorize(parameterVectorLinear=linearModel_dsd,
                             parameterVectorSigmoidal=sigmoidalModel_dsd,
                             parameterVectorDoubleSigmoidal=doubleSigmoidalModel_dsd)

print(outputCluster_dsd) # This should give double sigmoidal 
```
